---
title: "United Way Food Insecurity Predictions"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: scroll
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Xremove state-level maps
#Xround hover output
#Xround predicted probability in table output
#definition of expenditure in the intro page 
#Xmake neighborhood dividers less prominent (white dividers less thick)
#Xcolor to rocket instead of viridis
#consistent color --> goodness color mapping
#(reverse it for the probability scale)
#Probability of being food SECURE
#Xhave GEOID in the table so can look up map locations in table

#Incorporating Shiny (interactive) components into flexdashboard:
#https://rmarkdown.rstudio.com/flexdashboard/shiny.html#simple_example
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(tigris)
library(sp)
library(tidycensus)
library(rgeos)#spatial
library(DT)#datatable
library(rmapshaper)
library(leaflet)
library(htmltools)
library(totalcensus)

acs <- read.csv("acs_predictions.csv")#read acs data with predictions

#round predictions
acs$lasso_bin_pred <- round(acs$lasso_bin_pred,2)
acs$lasso_pred <- round(acs$lasso_pred,2)

#for merging
acs$GEOID <- as.character(paste0(acs$GEOID, substr(acs$X, 13, 13)))

#this is block groups w/in tracts
ia_shp <- readRDS("ia_shp.RDS")#block_groups(state = "IA")  #TIME INTENSIVE 1st run (so each dashboard run)

county_list <- readRDS("county_list.RDS")#unique(counties("Iowa")$NAME)#TIME INTENSIVE 1st run (so each dashboard run)
county_list <- county_list[order(county_list)]
all_counties <-   block_groups(state = 'IA', county = county_list) #also time intensive...

ia_shp_join <- left_join(ia_shp, acs, by="GEOID" ) %>%
  rmapshaper::ms_simplify( keep = 0.01, keep_shapes = TRUE)

ia_shp_join$COUNTYNAME <- convert_fips_to_names(ia_shp_join$COUNTYFP, states = rep("IA", nrow(ia_shp_join)), geo_header="COUNTY")

ia_shp_join$labels <- paste0(
  "<strong> County: </strong> ",
  ia_shp_join$COUNTYNAME, "<br/> ",
  "<strong> GEOID: </strong> ",
  ia_shp_join$GEOID, "<br/> ",
  "<strong> Food Expenditure: </strong> $",
  ia_shp_join$lasso_pred, "<br/> ",
  "<strong> Food Insecurity: </strong> ",
  ia_shp_join$lasso_bin_pred, "<br/> "
) %>%
  lapply(htmltools::HTML)

#for color scale
limit <- c(min(acs$lasso_pred,na.rm=TRUE), max(acs$lasso_pred,na.rm=TRUE))
limit_bin <- c(min(acs$lasso_bin_pred,na.rm=TRUE), max(acs$lasso_bin_pred,na.rm=TRUE))

pal_bin <- colorBin(
  palette = "YlOrRd", domain = ia_shp_join$lasso_bin_pred,
  bins = seq(0, max(ia_shp_join$lasso_bin_pred, na.rm = TRUE) + .01, by = .05)
)

pal <- colorBin(
  palette = "YlOrRd", domain = ia_shp_join$lasso_pred,
  bins = seq(floor(min(ia_shp_join$lasso_pred, na.rm = TRUE)), floor(max(ia_shp_join$lasso_pred, na.rm = TRUE)) + .1, by = 5)
)

```

Presence of Food Insecurity
===================================== 

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r, echo=FALSE}
selectInput("county", label = "Select County:", choices = unique(ia_shp_join$COUNTYNAME), selected = "Polk County")

```


Row 
-----------------------------------------------------------------------

### Introduction

The estimates refer to the proportion of food insecure households in each block group. The predictions are made in two stages. First, we use Current Population Survey data to estimate the probability that a given household is food insecure in at least one dimension in the past year. The LASSO technique is used to select demographic variables that improve out-of-sample prediction performance and survey sampling weights are used to decrease prediction bias. Second, we apply the first-stage parameter estimates to block-group data from the American Community Survey (ACS) to obtain the predicted proportion of food insecure households in each block group.



Row 
-----------------------------------------------------------------------


### County-level

```{r, echo=FALSE}
# renderPlot({ggplot(aes(fill  = lasso_bin_pred, colour=lasso_bin_pred),data = cnty_shp_join()) +
#   geom_sf()+
#   scale_fill_viridis_c("Predicted \nProbability",limits = limit_bin)+
#   scale_colour_viridis_c("Predicted \nProbability",limits = limit_bin)})


cnty_shp_join <- reactive({
  subset(ia_shp_join, COUNTYNAME == input$county) 
  #block_groups(state = 'IA', county = input$county)
  })
renderLeaflet({
leaflet(cnty_shp_join(),height=500, width=1000) %>%
  addTiles() %>%
  #setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal_bin(lasso_bin_pred),
    color = "white",
        stroke=FALSE,
    fillOpacity = 0.6,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  )%>%
  leaflet::addLegend(
    pal = pal_bin, values = ~lasso_bin_pred,
    opacity = 0.7, title = "Predicted \nProbability",
    position = "topright"
  )
})

```


Row 
-----------------------------------------------------------------------

### Predictions

```{r, echo=FALSE}
temp1 <- ia_shp_join %>% as.data.frame%>%dplyr::select(c("GEOID","lasso_bin_pred", "female", "avg_hhsize", "hispanic",
                                  "black", "kids", "elderly", "education", "employed",
                                  "married", "disability"))%>%
                  rename( 
                         `Predicted Probability` = lasso_bin_pred)
renderDataTable({
rbind(temp1[which(ia_shp_join$COUNTYNAME == input$county),],
      temp1[which(ia_shp_join$COUNTYNAME != input$county),])
  
})
```


Food Expenditure
===================================== 

Inputs {.sidebar}
-----------------------------------------------------------------------


```{r, echo=FALSE}
selectInput("county2", label = "Select County:", choices = unique(ia_shp_join$COUNTYNAME), selected = "Polk County")

```

Row 
-----------------------------------------------------------------------

### Introduction

The estimates refer to weekly average food expenditure per person. The predictions are made in two stages. First, we use Current Population Survey data to estimate weekly food expenditure per person at the household level. The LASSO technique is used to select demographic variables that improve out-of-sample prediction performance and survey sampling weights are used to decrease prediction bias. Second, we apply the first-stage parameter estimates to block-group data from the American Community Survey (ACS) to obtain average food expenditure per person in each block group.


Row 
-----------------------------------------------------------------------

### County-level
```{r,echo=FALSE}

cnty_shp_join2 <- reactive({
  subset(ia_shp_join, COUNTYNAME == input$county2) 
  })
renderLeaflet({
leaflet(cnty_shp_join2(),height=500, width=1000) %>%
  addTiles() %>%
  #setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal(lasso_pred),
    stroke=FALSE,
    color = "white",
    fillOpacity = 0.6,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  )%>%
  leaflet::addLegend(
    pal = pal, values = ~lasso_pred,
    opacity = 0.7, title = "Predicted \nFood Expenditure",
    position = "topright"
  )
})


```


Row 
-----------------------------------------------------------------------

### Predictions

```{r, echo=FALSE}
temp2 <- ia_shp_join %>%as.data.frame%>% dplyr::select(c("GEOID", "lasso_pred","female", "avg_hhsize", "hispanic",
                                  "black", "kids", "elderly", "education", "employed",
                                  "married", "disability"))%>%
                  rename(
                         `Predicted Expenditure` = lasso_pred)
renderDataTable({
rbind(temp2[which(ia_shp_join$COUNTYNAME == input$county2),],
      temp2[which(ia_shp_join$COUNTYNAME != input$county2),])
  
})
```

