---
title: "United Way Food Insecurity Predictions"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: scroll
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Incorporating Shiny (interactive) components into flexdashboard:
#https://rmarkdown.rstudio.com/flexdashboard/shiny.html#simple_example
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(tigris)
library(sp)
library(tidycensus)
library(rgeos)#spatial
library(DT)#datatable
library(rmapshaper)
library(leaflet)
library(htmltools)
library(totalcensus)

acs <- read.csv("acs_predictions.csv")#read acs data with predictions


#for merging
acs$GEOID <- as.character(paste0(acs$GEOID, substr(acs$X, 13, 13)))
#this is block groups w/in tracts
ia_shp <- readRDS("ia_shp.RDS")#block_groups(state = "IA")  #TIME INTENSIVE 1st run (so each dashboard run)

county_list <- readRDS("county_list.RDS")#unique(counties("Iowa")$NAME)#TIME INTENSIVE 1st run (so each dashboard run)
county_list <- county_list[order(county_list)]
all_counties <-   block_groups(state = 'IA', county = county_list) #also time intensive...

ia_shp_join <- left_join(ia_shp, acs, by="GEOID" ) %>%
  rmapshaper::ms_simplify( keep = 0.01, keep_shapes = TRUE)

ia_shp_join$COUNTYNAME <- convert_fips_to_names(ia_shp_join$COUNTYFP, states = rep("IA", nrow(ia_shp_join)), geo_header="COUNTY")

ia_shp_join$labels <- paste0(
  "<strong> County: </strong> ",
  ia_shp_join$COUNTYNAME, "<br/> ",
  "<strong> Food Expenditure: </strong> ",
  ia_shp_join$lasso_pred, "<br/> ",
  "<strong> Food Insecurity: </strong> ",
  ia_shp_join$lasso_bin_pred, "<br/> "
) %>%
  lapply(htmltools::HTML)

#for color scale
limit <- c(min(acs$lasso_pred,na.rm=TRUE), max(acs$lasso_pred,na.rm=TRUE))
limit_bin <- c(min(acs$lasso_bin_pred,na.rm=TRUE), max(acs$lasso_bin_pred,na.rm=TRUE))

pal_bin <- colorBin(
  palette = "viridis", domain = ia_shp_join$lasso_bin_pred,
  bins = seq(0, max(ia_shp_join$lasso_bin_pred, na.rm = TRUE) + .01, by = .05)
)

pal <- colorBin(
  palette = "viridis", domain = ia_shp_join$lasso_pred,
  bins = seq(floor(min(ia_shp_join$lasso_pred, na.rm = TRUE)), floor(max(ia_shp_join$lasso_pred, na.rm = TRUE)) + .1, by = 5)
)

```

Presence of Food Insecurity
===================================== 

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r, echo=FALSE}
selectInput("county", label = "Select County:", choices = unique(ia_shp_join$COUNTYNAME), selected = "Polk County")

```


Row 
-----------------------------------------------------------------------

### Introduction

Predictions are based on a logistic regression predicting the probability of a household experiencing some level of food insecurity. The LASSO technique is used to select demographic variables that improve out of sample prediction performance. Survey sampling weights are used to decrease prediction bias. 


Row 
-----------------------------------------------------------------------


### State-level
```{r, echo=FALSE}
cnty_shp_join <- reactive({
  subset(ia_shp_join, COUNTYNAME == input$county) 
  #block_groups(state = 'IA', county = input$county)
  })


#ggplot(aes(fill  = lasso_bin_pred, colour=lasso_bin_pred),data = ia_shp_join) +
#  geom_sf()+
#  scale_fill_viridis_c("Predicted \nProbability",limits = limit_bin)+
#  scale_colour_viridis_c("Predicted \nProbability",limits = limit_bin)
renderLeaflet({
leaflet(ia_shp_join,height=500, width=1000) %>%
  addTiles() %>%
  #setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal_bin(lasso_bin_pred),
    color = "white",
    fillOpacity = 0.6,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    ))
})
```

### County-level

```{r, echo=FALSE}
# renderPlot({ggplot(aes(fill  = lasso_bin_pred, colour=lasso_bin_pred),data = cnty_shp_join()) +
#   geom_sf()+
#   scale_fill_viridis_c("Predicted \nProbability",limits = limit_bin)+
#   scale_colour_viridis_c("Predicted \nProbability",limits = limit_bin)})

renderLeaflet({
leaflet(cnty_shp_join(),height=500, width=1000) %>%
  addTiles() %>%
  #setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal_bin(lasso_bin_pred),
    color = "white",
    fillOpacity = 0.6,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  )%>%
  leaflet::addLegend(
    pal = pal, values = ~lasso_bin_pred,
    opacity = 0.7, title = "Predicted \nProbability",
    position = "topright"
  )
})

```


Row 
-----------------------------------------------------------------------

### Predictions

```{r, echo=FALSE}
temp1 <- ia_shp_join %>% as.data.frame%>%dplyr::select(c("X","lasso_bin_pred", "female", "avg_hhsize", "hispanic",
                                  "black", "kids", "elderly", "education", "employed",
                                  "married", "disability"))%>%
                  rename(Location = X, 
                         `Predicted Probability` = lasso_bin_pred)
renderDataTable({
rbind(temp1[which(ia_shp_join$COUNTYFP == input$county),],
      temp1[which(ia_shp_join$COUNTYFP != input$county),])
  
})
```


Food Expenditure
===================================== 

Inputs {.sidebar}
-----------------------------------------------------------------------


```{r, echo=FALSE}
selectInput("county2", label = "Select County FIPS:", choices = unique(all_counties$COUNTYFP), selected = "111")

```

Row 
-----------------------------------------------------------------------

### Introduction

Predictions are based on a linear regression predicting the mean household food expenditure. The LASSO technique is used to select demographic variables that improve out of sample prediction performance. Survey sampling weights are used to decrease prediction bias.  


Row 
-----------------------------------------------------------------------

###  State-level
```{r,echo=FALSE}
cnty_shp_join2 <- reactive({
  subset(ia_shp_join, COUNTYFP == input$county2) 
  })
ggplot(aes(fill  = lasso_pred, colour=lasso_pred),data = ia_shp_join) +
  geom_sf()+
  scale_fill_viridis_c("Predicted \nExpenditure",limits = limit)+
  scale_colour_viridis_c("Predicted \nExpenditure",limits = limit)
```

### County-level
```{r,echo=FALSE}
renderPlot({
ggplot(aes(fill  = lasso_pred, colour=lasso_pred),data = cnty_shp_join2()) +
  geom_sf()+
  scale_fill_viridis_c("Predicted \nExpenditure",limits = limit)+
  scale_colour_viridis_c("Predicted \nExpenditure",limits = limit)})

```


Row 
-----------------------------------------------------------------------

### Predictions

```{r, echo=FALSE}
temp2 <- ia_shp_join %>%as.data.frame%>% dplyr::select(c("X", "lasso_pred","female", "avg_hhsize", "hispanic",
                                  "black", "kids", "elderly", "education", "employed",
                                  "married", "disability"))%>%
                  rename(Location = X, 
                         `Predicted Expenditure` = lasso_pred)
renderDataTable({
rbind(temp2[which(ia_shp_join$COUNTYFP == input$county2),],
      temp2[which(ia_shp_join$COUNTYFP != input$county2),])
  
})
```

